---
title: "Data Analysis - Chess Tournament"
author: "Trang Do"
date: '`r Sys.Date()`'
output:
  rmdformats::robobook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## INTRODUCTION
ELO chess rating system originated by is one of methods to the relative skill levels, expected score of players in two-player games such as Chess, Go, Draughts... This project is reading players' information from a text file for future of ELO chess rating system application

## 1. Preparing Data from a provided CSV

This part reads data of players in CSV file to dataframe in R. It also remove unnecessary data from CSV file

Reading CSV and remove redundant data
```{r library}
library(dplyr)
library(stringr)

path <- "https://raw.githubusercontent.com/Trangmis/607/master/project1/tournamentinfo.txt"
txtfile <- read.csv(path, header = F, sep = "|")

txtfile <- txtfile %>% filter(!(str_detect(V1,"--|Num|Pair")))
```

## 2. Converting Data to data frame

```{r}
oddrows <- txtfile %>% filter(row_number() %% 2 == 1) %>%
                      mutate(player = as.integer(V1),
                             playername = V2,
                             points = V3,
                             r1 = substr(V4,1,1), 
                             r1_opponent = as.integer(substr(V4,2,str_length(V4))),
                             r2 = substr(V5,1,1), 
                             r2_opponent = as.integer(substr(V5,2,str_length(V5))),
                             r3 = substr(V6,1,1), 
                             r3_opponent = as.integer(substr(V6,2,str_length(V6))),
                             r4 = substr(V7,1,1), 
                             r4_opponent = as.integer(substr(V7,2,str_length(V7))),
                             r5 = substr(V8,1,1), 
                             r5_opponent = as.integer(substr(V8,2,str_length(V8))),
                             r6 = substr(V9,1,1), 
                             r6_opponent = as.integer(substr(V9,2,str_length(V9))),
                             r7 = substr(V10,1,1), 
                             r7_opponent = as.integer(substr(V10,2,str_length(V10)))) 
    
evenrows <- txtfile %>% 
          filter(row_number() %% 2 == 0)%>%
          mutate(player = row_number(),
             uscf_id = as.integer(substr(V2,1,str_locate_all(pattern = " / R:", V2)[[1]][1])),
             startp = str_locate(pattern = ":", V2)[[1]][1]+1,
             end1 = str_locate(V2,"P")-1,
             end2 = str_locate(V2,">")-2
             )
evenrows$end1[is.na(evenrows$end1)] <- 0

evenrows <- evenrows %>%
                  mutate(pre_rating=as.integer(substr(V2,startp,ifelse(end1==0,end2,end1))))

evenrows <- evenrows %>% select (player,uscf_id , pre_rating)
oddrows <- oddrows %>% select (player,playername, points, r1,r1_opponent,
                                r2,r2_opponent, r3,r3_opponent,
                               r4,r4_opponent, r5,r5_opponent,
                               r6,r6_opponent, r7,r7_opponent)

df <- merge(evenrows,oddrows,by="player")
head(df)
```

## 3. PostgreSQL

### Calculating player's average/expected score & Writing data frame to PostgreSQL

```{r}
library(DBI)

db <- 'project1'  #provide the name of your db
host_db <- 'localhost'
db_port <- '5432'
db_user <- 'ass2' 
db_password <- 'data607'

con <- dbConnect(RPostgres::Postgres(), dbname = db, host=host_db, port=db_port, user=db_user, password=db_password) 

dbWriteTable(con, "df", df, overwrite = TRUE, row.names = FALSE)

sqlresult <- "SELECT df_games_1.player,allplayed_unplayed.playername, 
            	 allplayed_unplayed.points,allplayed_unplayed.pre_rating,round,
            	 df_games_1.status,df.playername as opponent_name, 
            	 df.points as opponent_points,
            	 df.pre_rating as opponent_pre_rating,
            	 AVG(df.pre_rating) OVER (PARTITION BY df_games_1.player), 
            	 (1/(1+10^((allplayed_unplayed.pre_rating-df.pre_rating)/400))) as expected_score
              FROM df, (SELECT player,
              		 UNNEST(
              			array [r1_opponent, r2_opponent,r3_opponent,r4_opponent, 
              			r5_opponent, r6_opponent,r7_opponent]) As opponent,
              		 UNNEST(ARRAY [r1, r2, r3, r4, r5, r6, r7]) AS status,
              		 UNNEST(ARRAY [1, 2, 3, 4, 5, 6, 7]) AS round
              		 FROM df) as df_games_1, (SELECT * FROM df
              		          WHERE (r1 in ('L','W','D') and r2 in ('L','W','D') and 
              			               r3 in ('L','W','D') and r4 in ('L','W','D') and
              			               r5 in ('L','W','D') and r6 in ('L','W','D') and 
              			               r7 in ('L','W','D'))
              		          OR 
                          			 (r1='U' or r2='U' or r3='U' or r4='U' or
                          				r5='U' or r6='U' or r7='U')) as allplayed_unplayed
               WHERE df_games_1.player = allplayed_unplayed.player and 
              	   df.player = df_games_1.opponent
              ORDER BY player,round"

dbquery <- dbGetQuery(con,sqlresult)
dbWriteTable(con, "project1_result", dbquery, overwrite = TRUE, row.names = FALSE)
dbDisconnect(con)
```

## 4. CSV File

### ELO Chess calculation an expected score in R data frame & writing to CSV file

```{r}
# Use dataframe in R
expectedscore <- dbquery %>% select (playername,opponent_name,round, status)%>%
                mutate(expected_score= 1/(1+10^((dbquery$pre_rating 
                                                 - dbquery$opponent_pre_rating)/400)))
# Write AVG to CSV file
write.csv(expectedscore,"Tdo_Project1_Export.csv")
```