---
title: "Data Analysis - Chess Tournament"
author: "Trang Do"
date: '`r Sys.Date()`'
output:
  rmdformats::robobook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Preparing Data from a provided CSV

This part reads data of players in CSV file to dataframe in R. It also unnecessary data from CSV file

Reading CSV and remove redundant data
```{r library}
library(dplyr)
library(stringr)

path <- "https://raw.githubusercontent.com/Trangmis/607/master/project1/tournamentinfo.txt"
txtfile <- read.csv(path, header = F, sep = "|")

txtfile <- txtfile %>% filter(!(str_detect(V1,"--|Num|Pair")))

```

## 2. Converting Data to data frame

** Initialize Data Frame
```{r initialize-data-frame}
players_frame <- data.frame(playerid=0,playername='',totalpoints=0,R=0,
                            averagepoint=0,predictedpoint=0,
                            r1_status='',r1_opponent=0,r2_status=0,r2_opponent=0,
                            r3_status='',r3_opponent=0,r4_status='',r4_opponent=0,
                            r5_status='',r5_opponent=0,r6_status='',r6_opponent=0,
                            r7_status='',r7_opponent=0)
players <- data.frame(playerid=0,playername='',totalpoints=0,r=0,
                            averagepoint=0,predictedpoint=0)
player_games <- data.frame(playerid=0, opponent=0,status='',round=0)
```

** Scan data from text file and write to dataframe

```{r write-data-to-dataframe}
for (i in 1:NROW(txtfile)){
  (i %% 2)
  if ((i%%2)>0){
    vplayer <- as.integer(txtfile$V1[i])
    r1status <- substr(txtfile$V4[i],1,1)
    r2status <- substr(txtfile$V5[i],1,1)
    r3status <- substr(txtfile$V6[i],1,1)
    r4status <- substr(txtfile$V7[i],1,1)
    r5status <- substr(txtfile$V8[i],1,1)
    r6status <- substr(txtfile$V9[i],1,1)
    r7status <- substr(txtfile$V10[i],1,1)
    r1o <- as.integer(substr(txtfile$V4[i],2,7))
    r2o <- as.integer(substr(txtfile$V5[i],2,7))
    r3o <- as.integer(substr(txtfile$V6[i],2,7))
    r4o <- as.integer(substr(txtfile$V7[i],2,7))
    r5o <- as.integer(substr(txtfile$V8[i],2,7))
    r6o <- as.integer(substr(txtfile$V9[i],2,7))
    r7o <- as.integer(substr(txtfile$V10[i],2,7))
    rpoint <- as.double(substr(txtfile$V2[i+1],15,19))
    
    newplayer <- data.frame(playerid=vplayer,playername=txtfile$V2[i],
                            totalpoints=as.double(txtfile$V3[i]),R=rpoint,
                            averagepoint=0.0,predictedpoint=0.0,
                            r1_status=r1status,r1_opponent=r1o,
                            r2_status=r2status,r2_opponent=r2o,
                            r3_status=r3status,r3_opponent=r3o,
                            r4_status=r4status,r4_opponent=r4o,
                            r5_status=r5status,r5_opponent=r5o,
                            r6_status=r6status,r6_opponent=r6o,
                            r7_status=r7status,r7_opponent=r7o)
    players_frame <-rbind(players_frame,newplayer)
    
    newplayer <- data.frame(playerid=vplayer,playername=txtfile$V2[i],
                            totalpoints=as.double(txtfile$V3[i]),r=rpoint,
                            averagepoint=0.0,predictedpoint=0.0)
    players <-rbind(players,newplayer)
   }
}
players_frame <- players_frame %>% filter(playerid>0)
players <- players %>% filter(playerid>0)
```

```{r}
#---------------Players-Games
temp <- as.data.frame(players_frame %>% 
                        select (playerid,opponent=r1_opponent,status=r1_status) %>%
                        mutate(round = 1))

player_games <- rbind(player_games,temp)

temp <- as.data.frame(players_frame %>% 
                        select (playerid,opponent=r2_opponent,status=r2_status) %>%
                        mutate(round = 2))

player_games <- rbind(player_games,temp)

temp <- as.data.frame(players_frame %>% 
                        select (playerid,opponent=r3_opponent,status=r3_status) %>%
                        mutate(round = 3))

player_games <- rbind(player_games,temp)

temp <- as.data.frame(players_frame %>% 
                        select (playerid,opponent=r4_opponent,status=r4_status) %>%
                        mutate(round = 4))

player_games <- rbind(player_games,temp)

temp <- as.data.frame(players_frame %>% 
                        select (playerid,opponent=r5_opponent,status=r5_status) %>%
                        mutate(round = 5))

player_games <- rbind(player_games,temp)

temp <- as.data.frame(players_frame %>% 
                        select (playerid,opponent=r6_opponent,status=r6_status) %>%
                        mutate(round = 6))

player_games <- rbind(player_games,temp)

temp <- as.data.frame(players_frame %>% 
                        select (playerid,opponent=r7_opponent,status=r7_status) %>%
                        mutate(round = 7))

player_games <- rbind(player_games,temp)

player_games <- player_games %>% filter(playerid>0) %>% arrange(playerid)
```

## 3. All Played Games & Unplayed Games with data frame in R & CSV

## All Played Games
```{r all-games-players}
library(stringr)

# count totalgames
count_Games <- player_games %>% 
  select(playerid) %>%
  filter(player_games$status %in% c('W','D','L'))%>%
  group_by(playerid)%>%
  summarise(totalgames=n()) %>%
  arrange(desc(totalgames),playerid)
# filter players who played all games (7 games)
all_game_players <- count_Games %>% filter(totalgames == 7)

# Get opponent info.
allgameplayer <- merge(all_game_players,player_games, by = "playerid")

opponent_inf <- players %>% mutate (opponent = playerid,opponentname=playername,
                            opponent_totalpoints=totalpoints,opponent_prer=r)

opponent_inf <- opponent_inf %>% select(opponent,opponentname,opponent_totalpoints,opponent_prer)

result <- merge(opponent_inf,allgameplayer, by="opponent")
result <- result %>% select(playerid,round,status,opponent,opponentname,
                            opponent_totalpoints,opponent_prer) %>%
                            arrange(playerid,result$round)
result<-merge(result,players,by="playerid")
resultallgames <- result %>% select(playerid,playername,totalpoints,r,round,status,
                            opponent,opponentname,opponent_totalpoints,
                            opponent_prer) %>% arrange(playerid,round) %>%
                            mutate(category = "Played all games")
```

## One or more unplayed games
```{r unplayed-game-players}
library(stringr)

# count totalgames
count_Games <- player_games %>% 
  select(playerid) %>%
  filter(player_games$status == 'U')%>%
  group_by(playerid)%>%
  summarise(totalgames=n()) %>%
  arrange(desc(totalgames),playerid)
# filter players who have one or more unplayed games
all_game_players <- count_Games %>% filter(totalgames >=1)

# Get opponent info.
allgameplayer <- merge(all_game_players,player_games, by = "playerid")

opponent_inf <- players %>% mutate (opponent = playerid,opponentname=playername,
                            opponent_totalpoints=totalpoints,opponent_prer=r)

opponent_inf <- opponent_inf %>% select(opponent,opponentname,opponent_totalpoints,opponent_prer)

result <- merge(opponent_inf,allgameplayer, by="opponent")
result <- result %>% select(playerid,round,status,opponent,opponentname,
                            opponent_totalpoints,opponent_prer) %>%
                            arrange(playerid,result$round)
result<-merge(result,players,by="playerid")
resultunplayedgames <- result %>% select(playerid,playername,totalpoints,r,round,status,
                            opponent,opponentname,opponent_totalpoints,
                            opponent_prer) %>% arrange(playerid,round)%>%
                            mutate(category = ">1 unplayed games")
```

## Export Data Frame to CSV
1. Merge all games players and one or more unplayed games players.
2. Mean function will not display decimals if elements of a calculation are integer, or initial is not a double for example 0.0

```{r combine-and-export}
exportcsv <- rbind(resultunplayedgames,resultallgames)
write.csv(exportcsv,"Tdo_Assignment3_Export.csv")

temp <- exportcsv %>%
            group_by(playerid)%>%
            summarise(player_AVG_pre_rating = mean(opponent_prer))
average_rating <-merge(temp,exportcsv, by ="playerid")
write.csv(average_rating,"tdo_project1_average_rating.csv")
```

## 4. All Played Games & Unplayed Games with data frame in R & PostgreSQL

```{r}
# Export to PostgreSQL
library(DBI)

db <- 'project1'  #provide the name of your db
host_db <- 'localhost'
db_port <- '5432'
db_user <- 'ass2' 
db_password <- 'data607'

con <- dbConnect(RPostgres::Postgres(), dbname = db, host=host_db, port=db_port, user=db_user, password=db_password) 

dbWriteTable(con, "players", players, overwrite = TRUE, row.names = FALSE)
dbWriteTable(con, "players_frame", players_frame, overwrite = TRUE, row.names = FALSE)
dbWriteTable(con, "player_games", player_games, overwrite = TRUE, row.names = FALSE)
## Using tables: players, player_games

sqlstatement <- "SELECT *, AVG(opponent_pre_rating) OVER (PARTITION BY players.playerid)
                FROM players,
              	(SELECT result1.*, playername AS opponent_name, r AS opponent_pre_rating,
              	        totalpoints AS opponent_totalpoints
              	  FROM (SELECT player_games.*, playedgames,unplayedgames FROM player_games 
              				, (SELECT player_games.playerid, sum(
              						CASE
              						   WHEN status in ('L','W','D') THEN 1
              						   ELSE 0	
              						END) AS playedgames, sum(
              						CASE
              						   WHEN status = 'U' THEN 1
              						   ELSE 0	
              						END) AS unplayedgames
              					FROM player_games
              					GROUP BY playerid
              					HAVING (sum(
              							CASE
              							   WHEN status in ('L','W','D') THEN 1
              							   ELSE 0	
              							END)=7 or  sum(
              							CASE
              							   WHEN status = 'U' THEN 1
              							   ELSE 0	
              							END)>=1)) AS filtergames
              					WHERE player_games.playerid = filtergames.playerid) AS result1 
              		LEFT JOIN players on result1.opponent  = players.playerid
              	ORDER BY playerid) AS filterplayer
              WHERE filterplayer.playerid = players.playerid"

dbquery <- dbGetQuery(con,sqlstatement)

dbWriteTable(con, "tdo_project1_average_rating", dbquery, overwrite = TRUE, row.names = FALSE)
                  
dbDisconnect(con)
```

## 5. ELO Calculation

```{r}
average_rating
```
