---
title: "DATA 607 ASSIGNMENT - SQL and R"
author: "Trang Do"
date: "`r Sys.Date()`"
output:
  rmdformats::material
---

# Introduction:

The assignment focus on creating an R dataframe from different sources such as SQL databases,  CSV files.

Due to an free account, a survey https://www.surveymonkey.com/r/D2FCKW2 was created on SurveyMonkey.com can't download responses in any format csv, excel.... So the result was duplicated by using Google Form https://forms.gle/m4wBvZBi3pyyP98N8. A CSV file is used to import data to PostgerSQL.

# Preparing Data
## Create database & data for PostgreSQL from csv file.

STEP 1: Create DB movies

CREATE DATABASE movies
    WITH 
    OWNER = postgres
    ENCODING = 'UTF8'
    LC_COLLATE = 'English_United States.1252'
    LC_CTYPE = 'English_United States.1252'
    TABLESPACE = pg_default
    CONNECTION LIMIT = -1;

STEP 2: Create table    
-- Table: public.ratedmovies

-- DROP TABLE public.ratedmovies;

CREATE TABLE IF NOT EXISTS public.ratedmovies
(
    "Timestamp" date,
    score "char",
    black_widow "char",
    f9_the_fast_saga "char",
    the_jungle_cruise "char",
    the_suicide_quad "char",
    "hitmans_wifes_bodyguard" "char",
    the_marksman "char"
);

ALTER TABLE public.ratedmovies
    OWNER to ass2;
 
STEP 3: Using import tool in PostgreSQL to import data from csv file to to use

*** One important note to create a column (a field) for a table in PostgreSQL is that the uppercase letter in the column name causes an "ERROR:  column "....." does not exist" when using Select syntax.

## Access data from PostgreSQL

```{r opendb-connect}
library(DBI)
library(dplyr)

db <- 'movies'  #provide the name of your db
host_db <- 'localhost'
db_port <- '5432'
db_user <- 'ass2' 
db_password <- 'data607'

con <- dbConnect(RPostgres::Postgres(), dbname = db, host=host_db, port=db_port, user=db_user, password=db_password) 

dbquery <- dbGetQuery(con,'SELECT * FROM ratedmovies')

dbquery$black_widow[dbquery$black_widow == 'N'] <- NA   
dbquery$f9_the_fast_saga [dbquery$f9_the_fast_saga == 'N'] <- NA  
dbquery$the_jungle_cruise [dbquery$the_jungle_cruise == 'N'] <- NA 
dbquery$the_suicide_quad [dbquery$the_suicide_quad  == 'N'] <- NA 
dbquery$hitmans_wifes_bodyguard [dbquery$hitmans_wifes_bodyguard  == 'N'] <- NA 
dbquery$the_marksmand [dbquery$the_marksman  == 'N'] <- NA 

glimpse(dbquery)
```
## Reading CSV file

```{r csvfile}
csvpath <- "https://raw.githubusercontent.com/Trangmis/607/master/favoritemovies.csv"
csvdata <- read.csv(file = csvpath)

glimpse(csvdata)
```

# Analyzing Data
## PostgreSQL
1. Create a view in PostgreSQL for ranking the most favorite movies based on rating = 5

2. Show the result

```{r postgresql}
mostfavorite <- dbGetQuery(con,'SELECT * FROM mostfavorite')
mostfavorite
```
# Notes
All works that are done on PostgreSQL can be found here https://github.com/Trangmis/607/blob/master/Assigment2%20work%20on%20PostgreSQL

