---
title: "Course 607 - Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Trang Do

August 29, 2020

This project is about reading, processing dataset & a subset.
The Breast Cancer Dataset is a subset which is posted at https://archive.ics.uci.edu/ml/datasets/Breast+Cancer.

For more information about this subset such as creators, donors, Data Dictionary .... https://archive.ics.uci.edu/ml/datasets/Breast+Cancer


## Load data to data frame
```{r }
library(tidyverse)

vBCancer = "https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer/breast-cancer.data"
BCancer = read.csv(file = vBCancer, header = FALSE)

```

breast-cancer.data isn't included column names. Therefore, column names are changed According to information in the file https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer/breast-cancer.names 

## Rename Data Columns
```{r }
colnames(BCancer) <-c("class", "age","menopause","tumor_size","inv_nodes","node_caps","deg_malig",
                      "breast","breast_quad","irradiat")

```

## Get a subset from original dataset
```{r }
mysubset <- subset(BCancer, menopause == "premeno"  &  breast_quad == "left_low",
                   select = c("age","menopause","tumor_size","inv_nodes","node_caps","deg_malig",
                              "breast","breast_quad"))

view(mysubset)

```


## INFORMATION

Original dataset 
```{r }
summary(BCancer)

```
Subset 
```{r }
summary(mysubset)
```

## WORKS CITED

https://archive.ics.uci.edu/ml/datasets/Breast+Cancer

Dias, José. (2009). Breast cancer diagnostic typologies by grade-of-membership fuzzy modeling. 129-133. Fluctuation free approximations are used everywhere matrix representation is directly or indirectly involved. The idea is simple: The matrix representation of a function operator, whose action on its operand is the multiplication by a function, is equivalent ...
You can also embed plots, for example:

