---
title: "Assignment 10"
author: "Trang Do"
date: "`r Sys.Date()`"
output: openintro::lab_report
---
# TEXT MINING IN R

This assignment is a continuation of the exploration and observation of the NYTimes Book APIs from assignment 9, posted at https://rpubs.com/Trangmis/826396. The book APIs provide reviews of a book in NY Times newspapers if there are any. This exercise uses Text mining in R and sentiments of NRC, AFINN & Bing to analyze reviews written in English.

Reviews need to break into individual tokens. It's called tokenization, and the function unnest_tokens(word, text) is applied in this process.

## Question 1: Are NRC, AFINN & BING lexicon similar?

Two books '21 LESSONS FOR THE 21ST CENTURY', 'ALEXANDER HAMILTON' are linked to NRC, AFINN & BING lexicon. Plots show that there is similarity across lexicons.

### Load packages & prepare data
```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(dplyr)
library(tidytext)
library(ngram)
library(DBI)
library(htmltools)
library(forcats)
library(reshape2)
library(gridExtra)

bing_dictionary <- get_sentiments("bing")
afinn_dictionary <- get_sentiments("afinn")
nrc_dictionary <- get_sentiments("nrc")

# ---- Codes to download reviews -------
db <- 'nytimes_books'
host_db <- 'localhost'
db_port <- '5432'
db_user <- 'postgres' 
db_password <- 'data607'

con <- dbConnect(RPostgres::Postgres(), dbname = db, host=host_db, port=db_port, user=db_user, password=db_password) 

#sqlstatement <-"SELECT
#                    DISTINCT ON (book_review_link) book_review_link, isbn10, title
#                FROM
#                   books
#              WHERE book_review_link !='
#              ORDER BY
#                  book_review_link"

#query <- dbGetQuery(con, sqlstatement)
#mycorpus <- data.frame()
#n<-lengths(query)[[1]]
#for (i in 1:n){
#  response<-read_html(query$book_review_link[[i]])
#  text_df <- tibble(isbn=query$isbn10[i],booktitle = query$title[i],
#                    line = 1:length(response), text = response)  
# test <- text_df%>% unnest_tokens(word,text)
#  mycorpus <- rbind(mycorpus,test)
#  Sys.sleep(15)
#}

#dbWriteTable(con,'mycorpus',mycorpus,overwrite=TRUE)
# --------------------------------------

mycorpus <- dbGetQuery(con, "select mycorpus.* from mycorpus")

mycorpus <- mycorpus %>% select(booktitle,word) %>%
            group_by(booktitle)%>%
            count(word, sort = TRUE) 

nrc.bookreview <- mycorpus %>% inner_join(nrc_dictionary,by="word")
afinn.bookreview <- mycorpus %>% inner_join(nrc_dictionary,by="word")
bing.bookreview <- mycorpus %>% inner_join(nrc_dictionary,by="word")
```

### Link text of reviews to NRC sentiments

The results of tokenization link to NRC Lexicon and plot for any analyses might have

```{r nrc-sentiments}
test <- nrc.bookreview %>%
  filter(booktitle %in% c("21 LESSONS FOR THE 21ST CENTURY","ALEXANDER HAMILTON")) %>%
  group_by(booktitle, sentiment)%>%
  summarise(freq=sum(n))
  
test %>%
  mutate(sentiment = fct_reorder(sentiment,freq)) %>%
  ggplot(aes(x=sentiment, y=freq,fill=sentiment)) +
    geom_bar(stat="identity", width=.4) +
    facet_wrap(~booktitle,ncol=2)+
    coord_flip() +
    xlab("Sentiment")+
    ylab("Count")+
    theme(legend.position="none")
```

### Link text of reviews to AFINN sentiments

```{r afinn-sentiments}
test <- afinn.bookreview %>%
  filter(booktitle %in% c("21 LESSONS FOR THE 21ST CENTURY","ALEXANDER HAMILTON")) %>%
  group_by(booktitle, sentiment)%>%
  summarise(freq=sum(n))

test %>%
  mutate(sentiment = fct_reorder(sentiment,freq)) %>%
  ggplot(aes(x=sentiment, y=freq,fill=sentiment)) +
    geom_bar(stat="identity", width=.4) +
    facet_wrap(~booktitle,ncol=2)+
    coord_flip() +
    xlab("Sentiment")+
    ylab("Count")+
    theme(legend.position="none")
```

### Link text of reviews to BING sentiments

```{r bing-sentiments}
test <- bing.bookreview %>%
  filter(booktitle %in% c("21 LESSONS FOR THE 21ST CENTURY","ALEXANDER HAMILTON")) %>%
  group_by(booktitle, sentiment)%>%
  summarise(freq=sum(n))

test %>%
  mutate(sentiment = fct_reorder(sentiment,freq)) %>%
  ggplot(aes(x=sentiment, y=freq,fill=sentiment)) +
    geom_bar(stat="identity", width=.4) +
    facet_wrap(~booktitle,ncol=2)+
    coord_flip() +
    xlab("Sentiment")+
    ylab("Count")+
    theme(legend.position="none")
```

## Question 2: Is review differ from others based on Books, Book Category?

Children books' reviews are more positive words than others.

```{r echo=FALSE}
mycorpus <- dbGetQuery(con, "select mycorpus.*,list_name 
                                from mycorpus,books
                                where mycorpus.isbn=books.isbn10")

test <- mycorpus %>%
  inner_join(get_sentiments("bing"), by="word") %>%
  filter(booktitle %in% c("RED NOTICE","THE PRINCESS DIARIST","21 LESSONS FOR THE 21ST CENTURY","ALEXANDER HAMILTON","BROWN GIRL DREAMING","CAPTAIN UNDERPANTS"))%>%
  count(list_name,booktitle, index = line, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate('Book_Category'=list_name,sentiment = positive - negative)

ggplot(test, 
       aes(index, sentiment, fill = Book_Category)) +
  geom_col() +
  facet_wrap(~booktitle, ncol = 2) + 
  theme(
    legend.position = "bottom"
    )
```

## For More Expansion
1. Does a book review match with a book?
2. Does context go wrong if text mining is only based on the individual token?
3. Can Lexicon learn a word from context, sentence(s) as humans learn a language?

## Work Cited
This dataset was published in Saif M. Mohammad and Peter Turney. (2013), ``Crowdsourcing a Word-Emotion Association Lexicon.'' Computational Intelligence, 29(3): 436-465.

article{mohammad13,
author = {Mohammad, Saif M. and Turney, Peter D.},
title = {Crowdsourcing a Word-Emotion Association Lexicon},
journal = {Computational Intelligence},
volume = {29},
number = {3},
pages = {436-465},
doi = {10.1111/j.1467-8640.2012.00460.x},
url = {https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1467-8640.2012.00460.x},
eprint = {https://onlinelibrary.wiley.com/doi/pdf/10.1111/j.1467-8640.2012.00460.x},
year = {2013}
}

Name: AFINN-111 
 URL: http://www2.imm.dtu.dk/pubdb/views/publication_details.php?id=6010 
 License: Open Database License (ODbL) v1.0 
 Size: 78 KB (cleaned 59 KB) 
 Download mechanism: https 


...

