---
title: "Assignment3"
author: "Trang Do"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
library(dplyr)
library(stringr)

csvpath <- "https://raw.githubusercontent.com/Trangmis/607/master/Assignment3/majors-list.csv"
csvdata <- read.csv(file = csvpath)

glimpse(csvdata)

filtered_major <- csvdata %>% filter(str_detect(Major,"DATA|STATISTIC"))
head(filtered_major)
```
## Question 2
Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry" 
[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
fruit <-c("bell pepper","bilberry","blackberry","blood orange",
"blueberry","cantaloupe","chili pepper","cloudberry",
"elderberry","lime","lychee","mulberry","olive","salal berry")
fruit

newformat <- str_c(fruit,sep=",",collapse = "\", \"")
newformat <- paste("c(\"",newformat,"\")")
newformat <- str_replace(newformat, " b", "b")
newformat <- str_replace(newformat, "y ", "y")
cat(newformat)
```

## Question 3

Describe, in words, what these expressions will match:
A. (.)\1\1 : any string contains three consecutive of \1.

```{r}
test <- c("u\1\1\1apple", "apple")
str_view(test, "(.)\1\1", match = TRUE)
```

B. "(.)(.)\\2\\1" : same letter before and after of 2 same letter
```{r}
test <- c("bell pepper","bilberry","blackberry","blood orange",
"blueberry","cantaloupe","chili pepper","cloudberry",
"elderberry","lime","lychee","mulberry","olive","salal berry")

str_view(test, "(.)(.)\\2\\1", match = TRUE)
```

C.  "(..)\1" : \1\1 repeats once time after the first appear of \1\1
```{r}
test <- c("bell pepper","bilberry","blackberry","blood orange",
"blueberry","cantaloupe","chili pepper","cloudberry",
"elderberry","lime","lychee","mulberry","olive","salal berry",
"\1\1\1\1apple")

str_view(test, "(..)\1", match = TRUE)
```

D. "(.).\\1.\\1": The letter is located that has any letter after it. 
Then the letter with any letter with the repeated letter 

```{r}
test <- c("bell pepper","bilberry","blackberry","blood orange",
"blueberry","cantaloupe","chili pepper","cloudberry",
"elderberry","lime","lychee","mulberry","olive","salal berry",
"mamamalakaya","momame")

str_view(test, "(.).\\1.\\1", match = TRUE)
```

D."(.)(.)(.).*\\3\\2\\1" 
    Ex : 3 letters follows by same three letters but the third swap the position with the first letter.
    
```{r}
test <- c("123123","123213","123231","123132","123321","123312",
"bebrbehg","banhanabem","dskegkgekek","thABCACBBCAn","thABCCBAn","thCBAABCn")

str_view(test, "(.)(.)(.).*\\3\\2\\1", match = TRUE)
```

## Question 4 Construct regular expressions to match words that:

A. Start and end with the same character
  "^(.)(.)*\\1$"
```{r}
test <- c("abba","sos","pump","test", "toast", "berry")

str_view(test, "^(.)(.)*\\1$", match = TRUE)
```     
B. Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
    "(..)(.)*\\1" : This regular expression will check any position of a string and its repeat (one repeat).
    "^(..)(.)*\\1" : This regular expression will check start of a string and its repeat (one repeat).
```{r}
test <- c("church","sos","pump","test", "toast", "berry","jhthjjstha")

str_view(test, "(..)(.)*\\1", match = TRUE)
str_view(test, "^(..)(.)*\\1", match = TRUE)
```  

C. Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
test <- c("eleven","possessiveness")
str_view(test, "(.)(.)(.).*\\1", match = TRUE)

```
 
 