---
title: "DATA 607-Assignment 5 - Resubmission"
author: "Trang Do"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## INTRODUCTION


## 1. Preparing Data from a provided CSV

```{r library, echo=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(kableExtra)

path <- "https://raw.githubusercontent.com/Trangmis/607/master/Assignment5/israeli_vaccination_data_analysis_start.csv"
csvfile <- read.csv(path,na.strings=c("",0))

csvfile[[3,1]] <-"<50"
csvfile[[5,1]] <-">50"
colnames(csvfile)<- c('Age','Not_Vax','Full_Vax',
                      'Severe_Not_Vax','Severe_Full_Vax','Efficacy')

csvfile <- csvfile %>% replace(is.na(.), 0)
```
## 2. Transpose Wide to Long Practice
```{r Transpose_data,echo = FALSE,results='asis'}

wide_data<- csvfile[(csvfile$Age == "<50" | csvfile$Age == ">50" ) &
                   !str_detect(csvfile$Not_Vax,"%"),]
wide_data <- rbind(wide_data,csvfile[(csvfile$Age == "<50" | csvfile$Age == ">50" ) &
                   str_detect(csvfile$Not_Vax,"%"),])

long_data <- gather(wide_data,key="Category","Value",2:6)

wide_data %>%
knitr::kable(caption="Data in Wide",digits = 2, 
               format.args = list(big.mark = ",", scientific = FALSE))%>%
  kable_styling(bootstrap_options = c("hover","bordered","striped","condensed"))%>%
  kable_paper(html_font="arial",font_size=10,full_width = F)

long_data %>%
knitr::kable(caption="Data in Long",digits = 2, 
               format.args = list(big.mark = ",", scientific = FALSE))%>%
  kable_styling(bootstrap_options = c("hover","bordered","striped","condensed"))%>%
  kable_paper(html_font="arial",font_size=10,full_width = F)
```

```{r Working_with_data, echo=FALSE}
vax_df <- csvfile[(csvfile$Age == "<50" | csvfile$Age == ">50" ) &
                   !str_detect(csvfile$Not_Vax,"%"),]

#Convert Text to number
vax_df$Not_Vax <- as.numeric(str_remove_all(vax_df$Not_Vax,","))
vax_df$Full_Vax <- as.numeric(str_remove_all(vax_df$Full_Vax,","))
vax_df$Severe_Not_Vax <- as.numeric(vax_df$Severe_Not_Vax)
vax_df$Severe_Full_Vax <- as.numeric(vax_df$Severe_Full_Vax)

percent_df <- csvfile[(csvfile$Age == "<50" | csvfile$Age == ">50" ) &
                   str_detect(csvfile$Not_Vax,"%"),]

percent_df$Not_Vax <- as.numeric(sub("%", "",percent_df$Not_Vax,fixed=TRUE))/100
percent_df$Full_Vax <- as.numeric(sub("%", "",percent_df$Full_Vax,fixed=TRUE))/100
percent_df$Severe_Not_Vax <- as.numeric(percent_df$Severe_Not_Vax)
percent_df$Severe_Full_Vax <- as.numeric(percent_df$Severe_Full_Vax)

vax_df %>% 
  knitr::kable(caption="Population",digits = 2,
               format.args = list(big.mark = ",",scientific = FALSE))%>%
  kable_styling(bootstrap_options = c("hover","bordered"))%>%
  kable_paper(html_font="arial",
              font_size=10,full_width = F)

percent_df %>% 
  knitr::kable(caption="Percentage",digits = 2,
               format.args = list(big.mark = ",",scientific = FALSE))%>%
  kable_styling(bootstrap_options = c("hover","bordered"))%>%
  kable_paper(html_font="arial",
              font_size=10,full_width = F)

wide <- vax_df %>% 
  full_join(percent_df, by ="Age")

wide$Severe_Not_Vax.y <- wide$Severe_Not_Vax.x*100000/wide$Not_Vax.x
wide$Severe_Full_Vax.y <- wide$Severe_Full_Vax.x*100000/wide$Full_Vax.x
wide$Efficacy.y <- (1 - wide$Severe_Full_Vax.y/wide$Severe_Not_Vax.y)

vax_df <- subset(wide %>%
                  mutate(Total_population = 
                           (Not_Vax.x + Full_Vax.x)/(Not_Vax.y+Full_Vax.y)),
                select=c(Age, Not_Vax.x,Full_Vax.x,Severe_Not_Vax.x,
                         Severe_Full_Vax.x,Total_population,Efficacy.x))
vax_df <- vax_df %>%
  rename( Not_Vax = Not_Vax.x, Full_Vax = Full_Vax.x,Severe_Not_Vax = Severe_Not_Vax.x,
          Severe_Full_Vax = Severe_Full_Vax.x, Efficacy = Efficacy.x)

percent_df <- subset(wide %>%
                  mutate(Total_population = 0),
                  select=c(Age, Not_Vax.y,Full_Vax.y,Severe_Not_Vax.y,
                         Severe_Full_Vax.y,Total_population,Efficacy.y))
percent_df <- percent_df %>%
  rename( Not_Vax = Not_Vax.y, Full_Vax = Full_Vax.y,Severe_Not_Vax = Severe_Not_Vax.y,
          Severe_Full_Vax = Severe_Full_Vax.y, Efficacy = Efficacy.y)

total <- vax_df %>%
        summarise(Age = "Total", Not_Vax=sum(Not_Vax),Full_Vax=sum(Full_Vax),
                  Severe_Not_Vax = sum(Severe_Not_Vax),
                  Severe_Full_Vax = sum(Severe_Full_Vax),
                  Total_population=0,Efficacy=0)

total [nrow(total ) + 1,] = c("Total",total$Not_Vax/sum(vax_df$Total_population),
                              total$Full_Vax/sum(vax_df$Total_population),
                              total$Severe_Not_Vax*100000/total$Not_Vax,
                              total$Severe_Full_Vax*100000/total$Full_Vax,0,
                              1-(total$Severe_Full_Vax/total$Full_Vax)/
                                (total$Severe_Not_Vax/total$Not_Vax))

total <- rbind(vax_df,percent_df,total)
total$Not_Vax <- as.numeric(total$Not_Vax)
total$Full_Vax <- as.numeric(total$Full_Vax)
total$Severe_Not_Vax <- as.numeric(total$Severe_Not_Vax)
total$Severe_Full_Vax <- as.numeric(total$Severe_Full_Vax)
total$Total_population <- as.numeric(total$Total_population)
total$Efficacy <- as.numeric(total$Efficacy)

total <- total %>% arrange(Age)
```

## Question 1:

What does this total population represent?

The total population is included people who are not vaccinated, fully vaccinated, or one dose of vaccine. The eligible age is 12 year old and above.

```{r Question_1 , echo=FALSE}
 total %>% filter (Age=='Total')%>% 
  knitr::kable(caption="Total Population",digits = 2,
               format.args = list(big.mark = ",",scientific = FALSE))%>%
  kable_styling(bootstrap_options = c("hover","bordered"))%>%
  kable_paper(html_font="arial",
              font_size=10,full_width = F)

```
# Question 2:

**Calculate the Efficacy vs. Disease; Explain your results.**

According to the result, the vaccine is more effective in the group 50 years old and below then the group 60 years old and above. 
```{r Question 2, echo = FALSE}
total %>%
  knitr::kable(caption="Survey Results",digits = 2,
               format.args = list(big.mark = ",",scientific = FALSE))%>%
  kable_styling(bootstrap_options = c("hover","bordered"))%>%
  kable_paper(html_font="arial",
              font_size=10,full_width = F)

```

## Question 3

From your calculation of efficacy vs. disease, are you able to compare the rate of severe cases in unvaccinated individuals to that in vaccinated 

The rate of severe cases in unvaccinated individuals are different to that in vaccinated by group of Age. Especialy, This data set shows a surprised difference of rate  of severe cases between unvaccinated an vaccinated individuals in the whole population to each group, and it might cause incorrect conclusion.
